---
title: "Model Performance on Forest Fire Dataset"
author: "Alison Jing Huang"
date: "4/15/2018"
output:
  pdf_document: default
  html_document: default
  word_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r echo=FALSE, message=FALSE}
library(ggplot2)
library(magrittr)
library(tidyverse)
library(knitr)
library(caret)
library(rpart)
library(devtools)
library(randomForest)
library(gridExtra)
```


## Load the data
```{r, echo=FALSE, message=FALSE}
df <- read.csv(file='/Users/Jing/Desktop/CSX415-Data-Science-and-Principles/csx415-project/project/ForestFire/data/forestfires.csv')
df$month <- as.numeric(df$month)
df$day <- as.numeric(df$day)
head(df)
```

After we conducted initial analysis on the dataset, the next step is to create a linear model on the first fires data. Recall earlier we have transform the raw datset to all numerical variables, and renamed it as `fires`, next use ggplot2 to examine different variable with respect to the response variable - **"AREA"**.


## Relationship between X and AREA
```{r}
df %>% 
ggplot() +
geom_bar(aes(x = X, y = area),
stat = 'identity') + 
labs(x="X", y="Burn Area") +
theme_bw()
```

## Relationship between Y and AREA
```{r}
df %>% 
ggplot() +
geom_bar(aes(x = Y, y = area), 
stat = 'identity') + 
labs(x="Y", y="Burn Area") +
theme_bw()
```

## Relationship between MONTH and AREA
```{r}
df %>% 
ggplot() +
geom_bar(aes(x = month, y = area), 
stat = 'identity') + 
labs(x="month", y="Burn Area") +
theme_bw()
``` 

## Relationship between DAY and AREA
```{r}
df %>% 
ggplot() +
geom_bar(aes(x = day, y = area), 
stat = 'identity') + 
labs(x="day", y="Burn Area") +
theme_bw()
``` 

## Relationship between FFMC and AREA
```{r}
df %>% 
ggplot() +
geom_bar(aes(x = FFMC, y = area), 
stat = 'identity') + 
labs(x="FFMC", y="Burn Area") +
theme_bw()
```

## Relationship between DMC and AREA
```{r}
df %>% 
ggplot() +
geom_bar(aes(x = DMC, y = area), 
stat = 'identity') + 
labs(x="DMC", y="Burn Area") +
theme_bw()
```

## Relationship between DC and AREA
```{r}
df %>% 
ggplot() +
geom_bar(aes(x = DC, y = area), 
stat = 'identity') + 
labs(x="DC", y="Burn Area") +
theme_bw()
```

## Relationship between ISI and AREA
```{r}
df %>% 
ggplot() +
geom_bar(aes(x = ISI, y = area), 
stat = 'identity') + 
labs(x="ISI", y="Burn Area") +
theme_bw()
```

## Relationship between TEMPERATURE and AREA
```{r}
df %>% 
ggplot() +
geom_bar(aes(x = temp, y = area), 
stat = 'identity') + 
labs(x="Temperature", y="Burn Area") +
theme_bw()
```

## Relationship between RH and AREA
```{r}
df %>% 
ggplot() +
geom_bar(aes(x = RH, y = area), 
stat = 'identity') + 
labs(x="RH", y="Burn Area") +
theme_bw()
```

## Relationship between wind and AREA
```{r}
df %>% 
ggplot() +
geom_bar(aes(x = wind, y = area), 
stat = 'identity') + 
labs(x="Wind", y="Burn Area") +
theme_bw()
```

## Relationship between rain and AREA
```{r}
df %>% 
ggplot() +
geom_bar(aes(x = rain, y = area), 
stat = 'identity') + 
labs(x="Rain", y="Burn Area") +
theme_bw()
```


# Model Performance

##1. Linear model on Variable "X"
```{r}
lm.x <- lm(df$area ~ df$X, data=df)
lm.x=lm(df$area~df$X)
lm.x
summary(lm.x)
layout(matrix(c(1,2,3,4),2,2))
g1 <-plot(lm.x)
```

##2.  Linear model on Variable "Y"
```{r}
lm.y <- lm(df$area ~ df$Y, data=df)
lm.y=lm(df$area~df$Y)
lm.y
summary(lm.y)
layout(matrix(c(1,2,3,4),2,2))
g2 <-plot(lm.y)
```

##3.  Linear model on Variable "month"
```{r}
lm.month <- lm(df$area ~ df$month, data=df)
lm.month=lm(df$area~df$month)
lm.month
summary(lm.month)
layout(matrix(c(1,2,3,4),2,2))
g3 <-plot(lm.month)
```

##4.  Linear model on Variable "day"
```{r}
lm.day <- lm(df$area ~ df$day, data=df)
lm.day=lm(df$area~df$day)
lm.day
summary(lm.day)
layout(matrix(c(1,2,3,4),2,2))
g4 <-plot(lm.day)
```

##5.  Linear model on Variable "FFMC"
```{r}
lm.FFMC <- lm(df$area ~ df$FFMC, data=df)
lm.FFMC=lm(df$area~df$FFMC)
lm.FFMC
summary(lm.FFMC)
layout(matrix(c(1,2,3,4),2,2))
g5 <-plot(lm.FFMC)
```

##6.  Linear model on Variable "DMC"
```{r}
lm.DMC <- lm(df$area ~ df$DMC, data=df)
lm.DMC=lm(df$area~df$DMC)
lm.DMC
summary(lm.DMC)
layout(matrix(c(1,2,3,4),2,2))
g6 <-plot(lm.DMC)
```

##7.  Linear model on Variable "DC"
```{r}
lm.DC <- lm(df$area ~ df$DC, data=df)
lm.DC=lm(df$area~df$DC)
lm.DC
summary(lm.DC)
layout(matrix(c(1,2,3,4),2,2))
g7 <-plot(lm.DC)
```

##8.  Linear model on Variable "ISI"
```{r}
lm.ISI <- lm(df$area ~ df$ISI, data=df)
lm.ISI=lm(df$area~df$ISI)
lm.ISI
summary(lm.ISI)
layout(matrix(c(1,2,3,4),2,2))
g8 <-plot(lm.ISI)
```

##9.  Linear model on Variable "temp"
```{r}
lm.temp <- lm(df$area ~ df$temp, data=df)
lm.temp=lm(df$area~df$temp)
lm.temp
summary(lm.temp)
layout(matrix(c(1,2,3,4),2,2))
g9 <-plot(lm.temp)
```

##10.  Linear model on Variable "RH"
```{r}
lm.RH <- lm(df$area ~ df$RH, data=df)
lm.RH=lm(df$area~df$RH)
lm.RH
summary(lm.RH)
layout(matrix(c(1,2,3,4),2,2))
g10 <-plot(lm.RH)
```

##11. Linear model on Variable "wind"
```{r}
lm.wind <- lm(df$area ~ df$wind, data=df)
lm.wind=lm(df$area~df$wind)
summary(lm.wind)
layout(matrix(c(1,2,3,4),2,2))
g11 <- plot(lm.wind)
```

##12. Linear model on Variable "rain"
```{r}
lm.rain <- lm(df$area ~ df$rain, data=df)
lm.rain=lm(df$area~df$rain)
lm.rain
summary(lm.rain)
layout(matrix(c(1,2,3,4),2,2))
g12 <- plot(lm.rain)
```

##13. Linear model including on the 12 attribute variables

```{r}
lm_all=lm(df$area~ df$X + df$Y + df$month + df$day + df$FFMC + df$DMC + df$DC + df$ISI + df$temp + df$RH + df$wind + df$rain, data = df)
lm_all
summary
layout(matrix(c(1,2,3,4),2,2))
g13<- plot(lm_all)
```





