---
title: "Linear Model Performance on Forest Fire Dataset"
author: "Alison Jing Huang"
date: "4/15/2018"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r echo=FALSE}
library(ggplot2)
library(magrittr)
library(tidyverse)
library(knitr)

```


##Load the data
```{r, echo=FALSE, message=FALSE}
df <- read.csv(file="https://archive.ics.uci.edu/ml/machine-learning-databases/forest-fires/forestfires.csv")
df$month <- as.numeric(df$month)
df$day <- as.numeric(df$day)
head(df)
```

After we conducted initial analysis on the dataset, the next step is to create a linear model on the first fires data. Recall earlier we have transform the raw datset to all numerical variables, and renamed it as `fires`, next use ggplot2 to examine different variable with respect to the response variable - **"AREA"**.


### Relationship between X and AREA
```{r load_packages, warning=FALSE, message=FALSE, echo=FALSE}
df %>% 
ggplot() +
geom_bar(aes(x = X, y = area),
stat = 'identity') + 
labs(x="X", y="Burn Area") +
theme_bw()

```

### Relationship between Y and AREA
```{r  message=FALSE, echo=FALSE}
df %>% 
ggplot() +
geom_bar(aes(x = Y, y = area), 
stat = 'identity') + 
labs(x="Y", y="Burn Area") +
theme_bw()
```

### Relationship between MONTH and AREA
```{r echo=FALSE, message=FALSE}
df %>% 
ggplot() +
geom_bar(aes(x = month, y = area), 
stat = 'identity') + 
labs(x="month", y="Burn Area") +
theme_bw()
``` 

### Relationship between DAY and AREA
```{r echo=FALSE, message=FALSE}
df %>% 
ggplot() +
geom_bar(aes(x = day, y = area), 
stat = 'identity') + 
labs(x="day", y="Burn Area") +
theme_bw()
``` 

### Relationship between FFMC and AREA
```{r echo=FALSE, message=FALSE}
df %>% 
ggplot() +
geom_bar(aes(x = FFMC, y = area), 
stat = 'identity') + 
labs(x="FFMC", y="Burn Area") +
theme_bw()
```

### Relationship between DMC and AREA
```{r echo=FALSE, message=FALSE}
df %>% 
ggplot() +
geom_bar(aes(x = DMC, y = area), 
stat = 'identity') + 
labs(x="DMC", y="Burn Area") +
theme_bw()
```

### Relationship between DC and AREA
```{r echo=FALSE, message=FALSE}
df %>% 
ggplot() +
geom_bar(aes(x = DC, y = area), 
stat = 'identity') + 
labs(x="DC", y="Burn Area") +
theme_bw()
```

### Relationship between ISI and AREA
```{r echo=FALSE, message=FALSE}
df %>% 
ggplot() +
geom_bar(aes(x = ISI, y = area), 
stat = 'identity') + 
labs(x="ISI", y="Burn Area") +
theme_bw()
```

### Relationship between TEMPERATURE and AREA
```{r echo=FALSE, message=FALSE}
df %>% 
ggplot() +
geom_bar(aes(x = temp, y = area), 
stat = 'identity') + 
labs(x="Temperature", y="Burn Area") +
theme_bw()
```

## Model Performance

### Linear model on Variable "X"
```{r echo=FALSE, message=FALSE}
lm.x <- lm(area ~ X, data=df)
lm.x=lm(area~X)
lm.x
summary(lm.x)
```

### Linear model on Variable "Y"
```{r echo=FALSE, message=FALSE}
lm.y <- lm(area ~ Y, data=df)
lm.y=lm(area~Y)
lm.y
summary(lm.y)
```