% !TEX encoding = UTF-8 Unicode
---
title: "Project Performance"
author: "Alison Jing Huang"
date: "4/21/2018"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE}
library(tidyverse)
library(dplyr)
library(kernlab)
library(ROCR)
library(psych)
library(ggplot2)
library(magrittr)
library(knitr)
```
### Load the data
```{r echo=FALSE, message=FALSE}
df <- read.table(file="~/Desktop/CSX415-Data-Science-and-Principles/csx415-project/ForestFire/data/forestfires.csv", sep = ",", header = T)
df$month <- as.numeric(df$month)
df$day <- as.numeric(df$day)
head(df)
```

# check the likelihood of Fire Observations
```{r echo=FALSE, message=FALSE}
round(table(df$area==0)/nrow(df),2)
```
The result shows that there is 48% percent of the time there is no observation of a forest fire.

## Linear Regression Models
###Boxplot of forest fire area for different X's
```{r echo=FALSE, message=FALSE}
par (mfrow=c(1,1))
df1 <- df[df$area>0,]
boxplot(log(area)~as.factor(X), data=df1, xlab="X", ylab="fire area", main ="Forest Fire Area for Different X's")
```

###Boxplot of forest fire area for different Y's
```{r echo=FALSE, message=FALSE}
boxplot(log(area)~as.factor(Y), data=df1, xlab="Y", ylab="fire area", main ="Forest Fire Area for Different Y's")
```

###Boxplot of forest fire area for different Months
```{r echo=FALSE, message=FALSE}
boxplot(log(area)~as.factor(month), data=df1, xlab="month", ylab="fire area", main ="Forest Fire Area for Different Months")
```

Here based on a total of 517 fire observations and distributions across different months, we can deduce that there are more fire observations in August and September.


###Boxplot of forest fire area for different days
```{r rcho=FALSE, message=FALSE}
boxplot(log(area)~day, data=df1, xlab="day", ylab="fire area", main="Forest Fire Area for Different Days")

reg_day <-lm(log(area)~ day, data= df1)
summary(reg_day)
```
###Scatterplots of all different predictor variables.
```{r echo=FALSE, message=FALSE}
par(mfrow =c(3,3))
plot(log(area) ~ FFMC + DMC + DC + ISI + temp + RH + wind + rain, data=df1)
par(mfrow=c(1,1))
```
