---
title: "More Data using Machine Learning Methods"
author: "Alison Jing Huang"
date: "5/6/2018"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
setwd("/Users/Jing/Desktop/CSX415-Data-Science-and-Principles/csx415-project/project/ForestFire")
library('ProjectTemplate')
load.project()
```

```{r echo=FALSE, message=FALSE}
library(ggplot2)
library(magrittr)
library(tidyverse)
library(knitr)
library(caret)
library(rpart)
library(rpart.plot)
library(devtools)
library(devtools)
library(randomForest)
library(tree)
library(RColorBrewer)
library(ROCR)
library(e1071)
```

### Load the data
```{r, echo=FALSE, message=FALSE}
df <- read.csv(file='/Users/Jing/Desktop/CSX415-Data-Science-and-Principles/csx415-project/project/ForestFire/data/forestfires.csv')
df$month <- as.numeric(df$month)
df$day <- as.numeric(df$day)
head(df)
```

### Create the training and test dataset with 70/30 Rule
```{r}
set.seed(517)
trainRowNumbers <- createDataPartition(df$area, p = 0.7, list = FALSE)
trainData <- df[trainRowNumbers,]
testData <- df[-trainRowNumbers,]

x = trainData[, 1:12]
y = trainData$area

```

### Validate 80/20 rule by checking size of training datset vs. test data
```{r}
# Dimension of trainning set
dim(trainData)
# Dimension of test set
dim(testData)
```

### Preprocessing & Trainning
First check to see the data contains any missing values by using anyNA() method.
"FALSE" indicates the the dataset contains all the values.
```{r rcho = FALSE, message= FALSE}
anyNA(df)
```

### 10-Fold Cross Validation 
```{r}
trctrl <- trainControl(method='repeatedcv', number = 10, verboseIter = TRUE)
set.seed(1034)
dtree_fit1 <- train(area ~., data= trainData, method ="rpart", parms = list(split = "information"), trControl = trctrl, tuneLength= 10)
dtree_fit1
plot(dtree_fit1)
```

### Random Forest Model Fitting with Training dataset:
```{r}
rf1 <- randomForest(area ~ X + Y + month + day + FFMC + DMC + DC + ISI + temp + RH + wind + rain, data=trainData)
print(rf1)
```

### Model Fitting with Testing dataset
```{r}
rf2 <- randomForest(area ~ X + Y + month + day + FFMC + DMC + DC + ISI + temp + RH + wind + rain, data=testData)
print(rf2)
```


### 5 X 5 Fold Cross-Validation
```{r}
trctrl <- trainControl(method='repeatedcv', number = 5, repeats = 5, verboseIter = TRUE)
set.seed(1034)
dtree_fit2 <- train(area ~., data= trainData, method ="rpart", parms = list(split = "information"), trControl = trctrl, tuneLength= 10)
dtree_fit2
plot(dtree_fit2)

```

### Recursive Feature Elimination(RFE) for choosing important features

```{r}
set.seed(1034)
options(warn=-1)

subsets <- c(1:12)
ctrl <- rfeControl(functions=rfFuncs, method='repeatedcv', repeats= 5, verbose = FALSE)

lmProfile <- rfe(x=trainData[, 1:12], y=trainData$area, size=subsets, rfeControl = ctrl)

lmProfile
```

Based on above RFE method result, the 5 most important features are identified as
**DMC**, **DC**, **temp**, **month** and **ISI**.


### Decision Tree Model

#### Model 1
1. Use all the factor variables with respect to the resposne varialble **AREA**
```{r}
tree_1 <- rpart(area~ X + Y + month + day + FFMC + DMC + DC + ISI + temp + RH + wind + rain, data = trainData)
tree_1$terms <- eval(tree_1$call$formula)
summary(tree_1)
rpart.plot(tree_1)
plotcp(tree_1)
```

#### Predictions from Tree Model 1
```{r}
testData <- testData[complete.cases(testData), ]
pred1 <- predict(tree_1, newdata = testData)
```


#### Model 2
2. Use the five important attribute variables namely **DMC**, **DC**, **temp**, **month** and **ISI** with respect to the response varialble **AREA**
```{r}
tree_2 <- rpart(area~ DMC + DC + temp + month + ISI, data = trainData)
tree_2$terms <- eval(model2$call$formula)
summary(tree_2)
rpart.plot(tree_2)
plotcp(tree_2)
```

#### Predictions from Tree Model 2
```{r}
predictions.tree_2 <- predict(tree_2, newdata = testData)
```

#### Calculate error = actual number of cases - predicted number of cases
```{r}
library(caret)
error <- testData$area - predictions.tree_1
sqrt(mean(error))
```

```{r}
error <- testData$area - predictions.tree_2
sqrt(mean(error))
```